Cosmological Evidences of Dark Matter through the CMB

### 2019-02-27 paper/main.tex
almost done

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{geometry}
 \geometry{
 a4paper,
 bottom=40mm
 %total={170mm,257mm},
 %left=20mm,
 %top=20mm,
 }
 \title{Cosmological Evidences of Dark Matter through the CMB}
\author{Lorenzo Speri}
\date{}
%\usepackage{hyperref}
\usepackage{natbib}
\usepackage{graphicx}
% ---- Maths -------
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{physics}
% ---- Links -------
%\usepackage{hyperref}
% citation color
\usepackage[colorlinks=true,citecolor=blue]{hyperref}
% for twnsore
\usepackage{tensind}
\tensordelimiter{:}

@@

\vspace{3cm}
\section{Introduction}
Many astronomical and cosmological observations suggest that, by using the General Theory of Relativity, there is an invisible form of matter that exerts a gravitational pull on baryonic matter. 
In order to explain such observations, scientists have proposed the existence of Dark Matter, a new form of matter that does not interact electromagnetically and, therefore, it appears invisible to us. 
The fundamental nature of Dark Matter is not well understood yet, and whether Dark Matter is truly a new form of matter or the consequence of new laws of gravity is still matter of debate.\\
One of the most compelling evidence for the existence of Dark Matter is given by the analysis of the Cosmic Microwave Background (CMB).\\
The Cosmic Microwave Background is the electromagnetic radiation coming from the hot plasma of the early stages of the Universe.
As the Universe was expanding, it cooled down to a temperature low enough to form stable hydrogen $e^- + p^+  \rightarrow H + \gamma$ (Recombination). 
As the number density of the free electrons dropped with the expansion, the Thomson scattering $e^- + \gamma  \rightarrow e^- + \gamma$ became inefficient and the photons have since streamed freely through the universe in all directions and they are today observed as the Cosmic Microwave Background \citep{LecturesPdf}.\\
The Cosmic Microwave Background appears to us as the best measured Planck spectrum that we know with fluctuations of order $10 ^{-5}$ around the average temperature $T_{CMB} =2.7$ K.
The fluctuations are extremely important because they contain the crucial information about the primordial plasma and its content. 
In fact, during the early stages of the Universe, the photons were interacting with baryonic matter, which in turn was interacting gravitationally with dark matter.
Thus, the observations of the CMB fluctuations are linked to the energy-matter content of our Universe and to dark matter.\\
The aim of this paper is to give a short and concise overview of the physics of the CMB and the cosmological evidence of Dark Matter through the analysis of the CMB fluctuations within the $\Lambda$CDM model assumptions.\\
We first give an historical overview of the discovery and measurements of the CMB, then we discuss the main assumptions of the $\Lambda$CDM model and the link between the content of our universe and the cosmological model. 
Then, the physics of the hot plasma is explained with a simplified treatment in order to explain the main features of the power spectrum of the CMB.
We conclude with the discussion of the CMB temperature power spectrum and the role of dark matter.

%The recent measurements of the Cosmic Microwave Background (CMB) radiation allowed us to infer the presence of Dark Matter in the Universe. In this summary we will explain qualitatively how the presence of Dark Matter influences the CMB anisotropies.
\pagebreak
\section{From the Discovery of the CMB to the Planck Mission}
%
%
%
\begin{figure}
\begin{center}
\includegraphics[width=0.7\textwidth]{Horn_Antenna.jpeg}
\caption{The 15 meter Holmdel horn antenna at Bell Telephone Laboratories in Holmdel, New Jersey.
(By NASA - Great Images in NASA Description, Public Domain, \url{https://commons.wikimedia.org/w/index.php?curid=6463768})}
\label{antenna}
\end{center}
\end{figure}
%
%
%
In the 1960s Penzias and Wilson measured an isotropic and constant with time strong signal coming from the sky by using a horn-reflector radio antenna (Figure(\ref{antenna})). 
They tried everything they could do to reduce the “noise” in their system, but the signal remained.\\
In 1965 Arno Penzias and Robert Wilson published the paper: \emph{A measurement of excess antenna temperature at
4080 Mc/s}, where they reported the measurements of an isotropic, unpolarized, and free from seasonal variations excess noise temperature
\citep{penziasMeasurementExcessAntenna1965}.
The explanation of that signal was given by Robert Dicke and his research group at Princeton University: such signal as the relic of an early, hot, dense, and opaque state of the universe\footnote{The existence of the cosmic background radiation had actually been predicted by the physicist George Gamow and his colleagues in 1948} \citep{bucherPhysicsCosmicMicrowave2015}
. \\
The discovery of the CMB created a big excitement and a new era of cosmology began.
Measuring the CMB spectrum and its deviation from the black body spectrum was the new challenge.
\\
First of all, the photons of the CMB can be absorbed by the Earth's atmosphere, because the energy per CMB photon (approximately $\sim 6 \times 10 ^{-4}$ eV) is comparable to the energy of vibration or rotation for a small molecule (of water for instance). 
In fact, microwaves with wavelengths shorter than $\lambda \sim 3$ cm are strongly absorbed by water molecules \citep{RydenIntroCosmoPdf}.
This problem could be overcomed by observing the CMB in different bands and from locations where the atmospheric humidity is low, at high altitudes and at low temperatures.
However, the best way to measure the CMB spectrum is to use satellites.\\
The first satellite that was launched to observe the CMB was COBE. 
It did provide a convincing first detection of the CMB anisotropy, and it played a crucial role in determining the viability of the different cosmological models at that time.
\\
After that, WMAP and PLanck space missions  increased the accuracy of the measurements Figure (\ref{cobe_wmap_planck}).\\
The COBE, WMAP and Planck missions measured the CMB anisotropies with increasing accuracy. COBE, the first CMB satellite, measured fluctuations to scales of 7$^\circ$. WMAP was able to measure resolutions down to 0.3$^\circ$ in five different frequency bands, with Planck measuring all the way down to just 5 arcminutes (0.07$^\circ$) in nine different frequency bands in total.\\
Nevertheless the satellites improved their accuracy, many other sources of noise were encountered such as the light coming from the center of our galaxy, other stars and other objects in the universe, in addition to the relative motion of the satellite with respect to the CMB (Figure (\ref{cobe_map})).\\ 
Finally, observations and statistical analysis showed that the CMB spectrum is close to a black body spectrum up to fluctuations of order of $10^{-5}$ K (Figure (\ref{cobe_blackbody})).\\
% maybe last sentence in the introduction with the image

\begin{figure}[h]
\centering
    \textbf{Cosmic Microwave Background  Spectrum}\par\medskip
\centering
   {\includegraphics[height=7cm]{blackbody}}
\caption{}
\label{cobe_blackbody}
\end{figure}

@@

\begin{figure}
\centering
    \textbf{The CMB sky measured by COBE}\par\medskip
\centering
\includegraphics[scale =0.3]{mono_di_cobe}
\caption{The microwave sky as seen by the COBE DMR (differential microwave radiometer) instrument. The top panel shows the microwave sky as seen on a linear temperature scale including zero. No anisotropies are visible in this image, because the CMB monopole at 2.725K dominates. In the middle panel, the monopole component has been subtracted. Apart from some slight contamination from the galaxy near the equator (corresponding to the plane of our Galaxy), one sees only a nearly perfect dipole pattern, owing to our peculiar motion with respect to the rest frame defined by the CMB. In the bottom panel, both the monopole and dipole components have been removed. Except for the galactic contamination around the equator, one sees the cosmic microwave background anisotropy along with some noise. (Credit: NASA/COBE Science Team)}
\label{cobe_map}
\end{figure}

@@

It is often said that our Universe is made up of $68.8$\% of Dark Energy, and $26.8 $\% by Dark Matter and $4.9 $\% of ordinary matter, but what do we really mean by that?\\
In this section we explain the meaning of these parameters and what is Universe made up of. 
The assumptions and the theoretical framework that we are going to explain will be then used to analyze the observations of the CMB.\\

@@

Another important assumption is that there exists a class of observers at rest with respect to the Cosmic Microwave Background \citep{bartelmannStandardModelCosmology}.

@@

Since Earth is not at rest with respect to the CMB, the satellite COBE measured the dipole pattern showed in the second panel in Figure (\ref{cobe_map}).\\
In addition we assume that space-time is described by a metric theory of gravity, and we treat space-time as a 4-dimensional manifold characterized by the metric tensor $g_{\mu \nu}$.

@@

where $k$ is the space curvature and it determines the geometry of the spatial dimensions, elliptical, Euclidean or hyperbolic for $k>0$, $k=0$ or $k<0$.

@@

Solving (\ref{einstein_eq})and using the above definition (\ref{energy-momentum-tensor}) (which is a rather long and tedious derivation), gives us the Friedmann equation:

@@

The first Friedmann equation tells us how the Hubble parameter$H(t)$, i.e. the expansion rate,  evolves according to what is contained in our Universe.

@@

Solving this for dust (a name for non-relativistic matter) $\rho_m c^2 << p_m$ and for radiation $\rho_r c^2 = p_r/3$ gives

@@

The energy density of radiation $u = \rho_r c^2 $ decreases with the power of four because photons experience not only the dilution due to the spatial expansion but their frequencies, accordingly their energies, decreased by the stretching of spacetime.\\

@@

We will see in the next sections how the CMB power spectrum puts constraints on these parameters and tell use that there is a form of matter that we cannot wee.

\section{Phenomenology of the Fluctuations}
The Cosmic Microwave Background fluctuations were produced when radiation and matter were interacting in the early stage of the Universe.
Before the release of the CMB at $z > 1100$ the electromagnetic radiation is firmly coupled (both energetically and kinetically) to the baryons.
The behavior of such a system can be solved by using the equations of relativistic hydrodynamics and the Einstein's field equations (\ref{einstein_eq}) with small perturbations of density $\delta \rho$, pressure $\delta p$ and gravitational potential  $\delta \Phi$.
Perturbations of different scales propagate as sound waves through the plasma until they are frozen at the time of recombination and they appear to us as temperature fluctuations $\Theta = \delta T /T$ in the CMB map.
For a photon gas we know that the temperature fluctuations are related to the number density $n$, energy density $u = \rho c^2$ and pressure $p$ as follow:
\begin{equation}
n \propto T ^3 \rightarrow \dfrac{\delta n}{n_0} = 3 \Theta \qquad \qquad \qquad
 u \propto T^4  \qquad p  = \dfrac{u}{3} \rightarrow \dfrac{\delta u}{u_0} = 4 \Theta = \dfrac{\delta p}{p_0} \notag
\end{equation}
Solving the full general-relativistic description of the photon-matter fluid requires a long and complicated treatment \cite{LecturesPdf}, so we use instead an analogy with the most favourite model of a physicist: the harmonic oscillator\footnote{As in every analogy in physics, the phenomena are not the same. A mass on a spring is not the same as a relativistic fluid perturbation in an expanding 4-dimensional manifold, but the analogy is very helpful for the understanding and the equations are not that different (in first approximation)}.\\
All kinds of perturbations are proportional to the temperature fluctuations $\Theta$, so we  can associate the fluctuations to the displacement $x$ of spring-mass system from its equilibrium.
Let us build the analogy.
The fluctuations of the primordial fluid can be characterized by a typical scale $\lambda = 2 \pi /k$ of oscillation/propagation that could depends on the local gravitational pull, pressure or other properties.
It is important to stress that $k$ is not a wavevector, but it behaves as it was. 
In fact, the $k$ parameter will represent the strength of the spring.
If $k$ is small, the spring is loose and it  oscillates over large scales $\lambda$, analogously the oscillations of the fluctuations will have a large wavelength. \\ 
In addition, fluctuations are damped due to the expansion of the Universe and due to collisions, and ultimately they feel a constant gravitational pull from the overall fluid.
By analogy with the mass-spring system we analyze the fluctuations at different scales by solving the well known equation:
\begin{equation}
m\ddot{x} = - k x - \eta \dot{x} - m g \qquad \rightarrow \qquad \ddot{x} + 2 \gamma \dot{x} + \omega_k ^2 x + g =0
\end{equation}
where $k$ is linked to the scale of oscillations $\omega_k ^2 = k/m$, $\eta$ to the damping ($2 \gamma = \eta /m$), $g$ to the constant gravitational pull of the overall system and $m$ the inertia of the perturbation.
The general solution of the differential equation is 
\begin{equation}
x = x_0 e^{i \, \omega t} - \dfrac{g \, m}{k} \qquad \qquad \omega = i \gamma \pm \sqrt{ \omega_k ^2 - \gamma^2}
\end{equation}
We will discuss the overall behavior of the photon-baryon fluid at different scales $\lambda$ by studying the solutions of the harmonic oscillator for different $k$.

\subsection{Sachs-Wolfe effect}
In the harmonic oscillator model for very small $k$ the mass experiences only the gravitational pull of the overdensities $x \approx - g m/k$ and it does not oscillate.\\
In the photon-baryon fluid, perturbation modes of very large scales (large $\lambda$, very small $k$) do not have the time to complete an entire oscillation before recombination and they are influenced only by the gravitational potential of the overdensities.
As a consequence of the Einstein's equivalence principle, photons that escape from a gravitational field are red-shifted.
If we consider the fluctuations of the CMB on large scales, the photons that tries to escape from the gravitational well created by matter are red-shifted according to the so-called Sachs-Wolfe effect. 
The theoretical result from the relativistic derivation gives a temperature fluctuation $\Theta = -\delta \Phi /(3c^2)$.
The Sachse Wolfe effect takes into account an additional effect that cannot be explained with the simple harmonic oscillator model and it arises from the time dilation due to the expansion of the Universe.

\subsection{Acoustic Oscillations and Damping}
\begin{wrapfigure}{r}{6cm}
\centering
%\begin{figure}
\includegraphics[scale=0.5]{baryonspring}
%\end{figure}
\end{wrapfigure}
Let us consider again the harmonic oscillator model in the case of $\omega_k > \gamma$, small scales where $k$ is big "enough".
The respective solutions are of the form of the damped harmonic oscillator with a shift of $-gm/k$
\[
x = x_0 e^{-\gamma t} e^{\pm i \sqrt{\omega_k ^2 - \gamma ^2} \, t} - \dfrac{g m}{k}
\]
Before recombination, the perturbations of the photon-baryon fluid behaves analogously to the damped harmonic oscillator.
The mass that elongates the spring and falls down is analogues to the fluid falling and compressing inside a gravitational potential.\\
During matter dominated era, overdensities of dark matter creates gravitational well where the photon baryon-fluid falls in.
As a consequence the gravitational potential compresses the baryon-photon fluid and the pressure starts to rise.
Since dark matter and baryonic matter are pressureless, the system accretes until the photon pressure causes the fluid to expand outward.\\
The process of compression and expansion gives rise to sound waves and it goes on until recombination.
Only the fluctuations that are caught at the extrema of their oscillations at recombination will become peaks in the CMB power spectrum.
Modes caught at the extrema form a harmonic series based on the distance sound can travel by recombination.
The first peak of the oscillations represents the mode that compressed the fluid inside potential wells before recombination only once, the second one represents the mode that compressed and then rarefied, the third one the mode that compressed then rarefied then compressed, etc. exactly as a spring!
The radiation pressure acts as a spring, whereas the inertia of the fluid as the mass of the ball.
Only the compression phase is influenced by the inertia of the fluid, i.e. if the mass is heavier the displacement will be bigger $-gm/k$.
\\
The temperature fluctuations are damped due to collisions between the baryons and the photons that dissipate energy, and the expansion of the universe.\\
%The damping scale provides another standard ruler for the curvature test
The physical scale of the damping depends on the baryon density through the mean free path
%And the matter density through the time available for the photons to random walk.

@@

\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{baryon_m}
\end{center}
\caption{aa}
\label{DM_bary}
\end{figure}

\section{Conclusion}
The many ways that baryons show up in the power spectrum imply that the power spectrum has many independent checks on the baryon density of the universe.  The baryon density is a quantity that the CMB can measure to exquisite precision.

>>>>

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{geometry}
 \geometry{
 a4paper,
 bottom=40mm
 %total={170mm,257mm},
 %left=20mm,
 %top=20mm,
 }
 \title{Cosmological Evidences of Dark Matter through the CMB}
\author{Lorenzo Speri}
\date{}
%\usepackage{hyperref}
\usepackage{natbib}
\usepackage{graphicx}
% ---- Maths -------
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{physics}
% ---- Links -------
%\usepackage{hyperref}
% citation color
\usepackage[colorlinks=true,citecolor=blue]{hyperref}
% for twnsore
\usepackage{tensind}
\tensordelimiter{:}
\usepackage{stackengine}

@@

\vspace{3cm}
\section{Introduction}
Many astronomical and cosmological observations suggest that, by using the General Theory of Relativity, there is an invisible form of matter that exerts a gravitational pull on baryonic matter. 
In order to explain such observations, scientists have proposed the existence of Dark Matter, a new form of matter that does not interact electromagnetically and, therefore, it appears invisible to us. 
The fundamental nature of Dark Matter is not well understood yet, and whether Dark Matter is truly a new form of matter or the consequence of new laws of gravity is still matter of debate.\\
One of the most compelling evidence for the existence of Dark Matter is given by the analysis of the Cosmic Microwave Background (CMB).\\
The Cosmic Microwave Background is the electromagnetic radiation coming from the hot plasma of the early stages of the Universe.
As the Universe was expanding, it cooled down to a temperature low enough to form stable hydrogen $e^- + p^+  \rightarrow H + \gamma$ (Recombination). 
As the number density of the free electrons dropped with the expansion, the Thomson scattering $e^- + \gamma  \rightarrow e^- + \gamma$ became inefficient and the photons have since streamed freely through the universe in all directions and they are today observed as the Cosmic Microwave Background \citep{LecturesPdf}.\\
The Cosmic Microwave Background appears to us as the best measured Planck spectrum that we know with fluctuations of order $10 ^{-5}$ around the average temperature $T_{CMB} =2.7$ K.
The fluctuations are extremely important because they contain the crucial information about the primordial plasma and its content. 
In fact, during the early stages of the Universe, the photons were interacting with baryonic matter, which in turn was interacting gravitationally with dark matter.
Thus, the observations of the CMB fluctuations are linked to the energy-matter content of our Universe and to dark matter.\\
The aim of this paper is to give a short and concise overview of the physics of the CMB and the cosmological evidence of Dark Matter through the analysis of the CMB fluctuations within the $\Lambda$CDM model assumptions.\\
We first give an historical overview of the discovery and measurements of the CMB, then we discuss the main assumptions of the $\Lambda$CDM model and the link between the energy-matter content of the Universe and the cosmological model. 
Afterwards, the physics of the primordial plasma is explained through an analogy with a spring-mass system.
In section 5 we explain how the phenomena in the primordial plasma are reflected in the measured temperature power spectrum of the CMB.
Lastly, we discuss how the different energy-matter components influence the CMB power spectrum and we conclude that dark matter is one of the leading constituents of the universe.

%The recent measurements of the Cosmic Microwave Background (CMB) radiation allowed us to infer the presence of Dark Matter in the Universe. In this summary we will explain qualitatively how the presence of Dark Matter influences the CMB anisotropies.
\pagebreak
\section{From the Discovery of the CMB to the Planck Mission}
%
%
%
\begin{figure}
\begin{center}
\includegraphics[width=0.7\textwidth]{Horn_Antenna.jpeg}
\caption{The 15 meter Holmdel horn antenna at Bell Telephone Laboratories in Holmdel, New Jersey.
(By NASA - Great Images in NASA Description, Public Domain, \url{https://commons.wikimedia.org/w/index.php?curid=6463768})}
\label{antenna}
\end{center}
\end{figure}
%
%
%
In the 1960s Penzias and Wilson measured an isotropic and constant with time strong signal coming from the sky by using a horn-reflector radio antenna shown in Figure(\ref{antenna}). 
They tried everything they could do to reduce the “noise” in their system, but the signal remained.\\
In 1965 Penzias and Wilson published the paper: \emph{A measurement of excess antenna temperature at
4080 Mc/s}, where they reported the measurements of an isotropic, unpolarized, and free from seasonal variations excess noise temperature
\citep{penziasMeasurementExcessAntenna1965}.
The explanation of that signal was given by Robert Dicke and his research group at Princeton University: Penzias and Wilson discovered the relic of an early, hot, dense, and opaque state of the universe\footnote{The existence of the cosmic background radiation had actually been predicted by the physicist George Gamow and his colleagues in 1948} \citep{bucherPhysicsCosmicMicrowave2015}
. \\
The discovery of the CMB created a big excitement and a new era of cosmology began.
Measuring the CMB spectrum and its deviation from the black body spectrum was the new challenge.
\\
First of all, the photons of the CMB can be absorbed by the Earth's atmosphere, because the energy per CMB photon (approximately $\sim 6 \times 10 ^{-4}$ eV) is comparable to the energy of vibration or rotation for a small molecule (of water for instance). 
In fact, microwaves with wavelengths shorter than $\lambda \sim 3$ cm are strongly absorbed by water molecules \citep{RydenIntroCosmoPdf}.
This problem could be overcome by observing the CMB in different bands and from locations where the atmospheric humidity is low, at high altitudes and at low temperatures.
However, any terrestrial experiment would be able to measure only a portion of the sky map and would be affected by Earth's atmosphere, so the best way to measure the CMB spectrum is to use satellites.\\
COBE was the first satellite launched to measure the CMB in 1989. 
It did provide a convincing first detection of the CMB anisotropy, and it played a crucial role in determining the viability of the different cosmological models at that time.
\\
After COBE, WMAP and Planck satellites measured the CMB anisotropies with increasing accuracy  Figure (\ref{cobe_wmap_planck}).\\
COBE measured fluctuations to scales of 7$^\circ$, WMAP was able to measure resolutions down to 0.3$^\circ$ in five different frequency bands, and Planck measured all the way down to just 5 arcminutes (0.07$^\circ$) in nine different frequency bands in total.\\
Nevertheless the satellites improved their accuracy, many other sources of noise were encountered such as the light coming from the center of our galaxy, other stars and other objects in the universe, in addition to the relative motion of the satellite with respect to the CMB (Figure (\ref{cobe_map})).\\ 
Finally, observations and statistical analysis showed that the CMB spectrum is close to a black body spectrum up to fluctuations of order of $10^{-5}$ K (Figure (\ref{cobe_blackbody})).\\
The technological efforts for precisely measuring the fluctuations of the CMB was a fundamental step-stone to understand what the universe is made of.
% maybe last sentence in the introduction with the image

\begin{figure}[h]
\centering
    \textbf{Cosmic Microwave Background  Spectrum}\par\medskip
\centering
   {\includegraphics[height=7cm]{blackbody}}
\caption{The monopole spectrum of the Cosmic Microwave Background Radiation using the original data set by the FIRAS team, available at \url{http://lambda.gsfc.nasa.gov/product/cobe/firas_monopole_get.cfm}. The vertical axis "MJy/sr" corresponds to $10^6$ jansky per steradian, where a jansky is $10^{-26}$ Watts per square-meter per Hertz. The horizontal axis ("1/cm") corresponds to the reciprocal of the microwave wavelength (in cm), which is proportional to the microwave frequency. The error bars are too small to be displayed by a computer screen, but vastly exaggerated error bars were included to show the measured data points.}
\label{cobe_blackbody}
\end{figure}

@@

\begin{figure}
\centering
    \textbf{The CMB sky measured by COBE}\par\medskip
\centering
\includegraphics[scale =0.3]{mono_di_cobe}
\caption{The microwave sky as seen by the COBE DMR (differential microwave radiometer) instrument. The top panel shows the microwave sky as seen on a linear temperature scale including zero. No anisotropies are visible in this image, because the CMB monopole at 2.725K dominates. In the middle panel, the monopole component has been subtracted. Apart from some slight contamination from the galaxy near the equator (corresponding to the plane of our Galaxy), one sees only a nearly perfect dipole pattern, owing to our peculiar motion with respect to the rest frame defined by the CMB. In the bottom panel, both the monopole and dipole components have been removed. Except for the galactic contamination around the equator, one sees the CMB anisotropy along with some noise \cite{bucherPhysicsCosmicMicrowave2015}. (Credit: NASA/COBE Science Team)}
\label{cobe_map}
\end{figure}

@@

It is often said that the Universe we live in is flat and it is made up of $69$\% of Dark Energy, and $26 $\% by Dark Matter and $5$\% of ordinary matter, but what do we really mean by that?\\
In this section we explain the meaning of these parameters and from which theoretical framework they come from. 
We are going to use the assumptions of the so-called $\Lambda$CDM model.\\

@@

Another important assumption is that there exists a class of observers at rest with respect to the Cosmic Microwave Background such that the previous assumptions are valid \citep{bartelmannStandardModelCosmology}.

@@

Since Earth is not at rest with respect to the CMB, the satellite COBE measured the dipole pattern showed in the second panel in Figure (\ref{cobe_map}).
And therefore from an observer on Earth the Universe is not isotropic.
In the following sections we consider only those observers which are at rest with respect to the Cosmic Microwave Background.\\
In addition we assume that spacetime is described by a metric theory of gravity, and we treat spacetime as a 4-dimensional manifold characterized by the metric tensor $g_{\mu \nu}$.

@@

By defining $g_{\mu \nu}$ we are basically  defining how time and space intervals are measured.
Spacetime is treated as a manifold because it shares the property with manifolds to be locally flat\footnote{locally flat means  that locally it appears as the Minkowski spacetime}.\\
$k$ is the space curvature and it determines the geometry of the spatial dimensions, elliptical, Euclidean or hyperbolic for $k>0$, $k=0$ or $k<0$.

@@

Solving (\ref{einstein_eq}) and using the above definition (\ref{energy-momentum-tensor}) (which is a rather long and tedious derivation), gives us the Friedmann equations:

@@

The first Friedmann equation tells us how the Hubble parameter $H(t)$, i.e. the expansion rate,  evolves according to what is contained in our Universe.

@@

Solving this for dust, i.e. non-relativistic matter, $\rho_m c^2 << p_m$ and for radiation $\rho_r c^2 = p_r/3$ gives

@@

The energy density of radiation $u = \rho_r c^2 $ decreases with the power of four because photons experience not only the dilution due to the spatial expansion but their frequencies, accordingly their energies, decrease by the stretching of spacetime.\\

@@

We will see in the next sections how the CMB power spectrum puts constraints on these parameters and tell us that there is a form of matter that we cannot see in $\Omega_{m0}$.

\section{Phenomenology of the Fluctuations}
The Cosmic Microwave Background fluctuations were produced when radiation and matter were interacting in the early stage of the Universe.
Before the release of the CMB at redshift $z > 1100$ the electromagnetic radiation is firmly coupled (both energetically and kinetically) to the baryons.
The behavior of such a system can be studied by solving the equations of relativistic hydrodynamics and the Einstein's field equations (\ref{einstein_eq}) with small perturbations of density $\delta \rho$, pressure $\delta p$ and gravitational potential  $\delta \Phi$.
Perturbations of different scales propagate as sound waves through the plasma until they are frozen at the time of recombination and they appear to us as temperature fluctuations $\Theta = \delta T /T$ in the CMB map.
For a photon gas we know that the temperature fluctuations are related to the number density $n$, energy density $u = \rho c^2$ and pressure $p$ as follow:
\begin{equation}
n \propto T ^3 \rightarrow \dfrac{\delta n}{n_0} = 3 \Theta \qquad \qquad \qquad
 u \propto T^4  \qquad p  = \dfrac{u}{3} \rightarrow \dfrac{\delta u}{u_0} = 4 \Theta = \dfrac{\delta p}{p_0} \notag
\end{equation}
Solving the full general-relativistic description of the photon-matter fluid requires a long and complicated treatment that can be found here \cite{LecturesPdf,keskitaloEffectMatterBaryon}. We use instead an analogy with the most favourite model of a physicist: the harmonic oscillator\footnote{As in every analogy in physics, the phenomena are not the same. A mass on a spring is not the same as a relativistic fluid perturbation in an expanding 4-dimensional manifold, but the analogy is very helpful for the understanding and the equations are not that different (in first approximation)}.\\
All kinds of perturbations are proportional to the temperature fluctuations $\Theta$, so we  can associate the fluctuations to the displacement $x$ of spring-mass system from its equilibrium.
Let us build the analogy.
The fluctuations of the primordial fluid can be characterized by a typical scale $\lambda = 2 \pi /k$ of oscillation/propagation that could depends on the local gravitational pull, pressure or other properties.
It is important to stress that $k$ is not a wavevector, but it behaves as it was. 
In fact, the $k$ parameter will represent the strength of the spring.
If $k$ is small, the spring is loose and it  oscillates over large scales $\lambda$, analogously the oscillations of the fluctuations will have a large wavelength. \\ 
In addition, fluctuations are damped due to the expansion of the Universe and due to collisions, and ultimately they feel a constant gravitational pull from the overall fluid.
By analogy with the mass-spring system we analyze the fluctuations at different scales by solving the well known equation:
\begin{equation}
m\ddot{x} = - k x - \eta \dot{x} - m g \qquad \rightarrow \qquad \ddot{x} + 2 \gamma \dot{x} + \omega_k ^2 x + g =0
\end{equation}
where $k$ is linked to the scale of oscillations $\omega_k ^2 = k/m$, $\eta$ to the damping ($2 \gamma = \eta /m$), $g$ to the constant gravitational pull of the overall system and $m$ the inertia of the perturbation.
The general solution of the differential equation is 
\begin{equation}
x = x_0 e^{i \, \omega t} - \dfrac{g \, m}{k} \qquad \qquad \omega = i \gamma \pm \sqrt{ \omega_k ^2 - \gamma^2}
\end{equation}
We will discuss the overall behavior of the photon-baryon fluid fluctuations with different scales $\lambda$ by studying the solutions of the damped harmonic oscillator for different $k$.

\subsection{Sachs-Wolfe effect}
In the harmonic oscillator model for very small $k$ the mass experiences only the gravitational pull of the overdensities $x \approx - g m/k$ and it does not oscillate.\\
In the photon-baryon fluid, perturbation modes of very large scales (large $\lambda$, very small $k$) do not have the time to complete an entire oscillation before recombination and they are influenced only by the gravitational potential of the overdensities.
As a consequence of the Einstein's equivalence principle, photons that escape from a gravitational field are red-shifted, photons that fall into a gravitational field are blue-shifted.
Photons associated with fluctuations with large $\lambda$ try to escape from the gravitational well created by matter at the time of recombination and, therefore, they are red-shifted.
A photon which happens to be at a potential maximum when the universe became transparent gains energy as it falls down the “potential hill”, and thus is blue-shifted.
A detailed general relativistic calculation gives a temperature fluctuation $\Theta = -\delta \Phi /(3c^2)$.
This is called the Sachs-Wolfe effect.
The Sachs-Wolfe effect takes into account an additional effect that cannot be explained with the simple harmonic oscillator model and it arises from the time dilation due to the expansion of the Universe.
%
%
%
%\begin{wrapfigure}{r}{4cm}

\begin{figure}
\centering
\includegraphics[scale=0.4]{baryonspring}
\caption{Analogy between the photon-baryon fluid and the spring-mass system. (Credit: \cite{huweb})}
\label{baryon_spring}
\end{figure}
%\end{wrapfigure}
%
%
%
\subsection{Acoustic Oscillations and Damping}

Let us consider the harmonic oscillator model in the case of $\omega_k > \gamma$, small scales where $k$ is big "enough".
The respective solutions are of the form of the damped harmonic oscillator with a shift of $-gm/k$
\[
x = x_0 e^{-\gamma t} e^{\pm i \sqrt{\omega_k ^2 - \gamma ^2} \, t} - \dfrac{g m}{k}
\]
Before recombination, the perturbations of the photon-baryon fluid behaves analogously to the damped harmonic oscillator.
The mass elongates the spring and it falls down as the photon-baryon fluid falls into a gravitational potential and it is compressed (Figure(\ref{baryon_spring})).\\
During matter dominated era, overdensities of dark matter creates gravitational well where the photon baryon-fluid falls in.
As a consequence, the gravitational potential compresses the baryon-photon fluid and the pressure starts to rise.
Since dark matter and baryonic matter are pressureless, the system accretes until the photon pressure causes the fluid to expand outward.\\
The process of compression and expansion gives rise to sound waves and it goes on until recombination.
The inward and outward oscillations of the photon-baryon fluid are called acoustic oscillations, since they represent a type of standing sound wave in the photon-baryon fluid\cite{RydenIntroCosmoPdf}.\\
Only the fluctuations that are caught at the extrema of their oscillations at recombination will become peaks in the CMB power spectrum.
Modes caught at the extrema form a harmonic series based on the distance sound can travel by recombination.
The first peak of the oscillations represents the mode that compressed the fluid inside potential wells before recombination only once, the second one represents the mode that compressed and then rarefied, the third one the mode that compressed then rarefied then compressed, etc. exactly as a spring!\\
The radiation pressure acts as a spring, whereas the inertia of the fluid as the mass of the ball.
Only the compression phase is influenced by the inertia of the fluid, i.e. if the mass is heavier the displacement will be bigger $-gm/k$.
\\
The particles of the photon-baryon fluid experience also collisions and this leads to a damping effect on the oscillations.
The physical scale of the damping depends on the baryon and photon density through the mean free path.
An additional damping effect comes from the expansion of the Universe.
%The damping scale provides another standard ruler for the curvature test

%And the matter density through the time available for the photons to random walk.

@@

\begin{figure}[h]
\begin{center}
\includegraphics[width=\textwidth]{baryon_m}
\end{center}
\caption{(Credit: \cite{huCosmicMicrowaveBackground2001})}
\label{DM_bary}
\end{figure}

\section{Conclusion}
The Planck Collaboration estimated the density parameters with incredible precision:

Substantially the curvature and the cosmological constant do not change the shape of the acoustic peaks but they only change shift their position on the power spectrum.

The many ways that baryons show up in the power spectrum imply that the power spectrum has many independent checks on the baryon density of the universe.  The baryon density is a quantity that the CMB can measure to exquisite precision.
